---
title: "MD Chapters 1-3"
author: "Kamran Acito"
format: html
editor: visual
---

## Question 1

```{r}
#| message: false
library(tidyr)
library(tidyverse)
```

```{r}
exercise_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/visualize_data.csv")
glimpse(exercise_data)

```

I expect there to by a negative relationship between exercise and BMI. When exercise increases, BMI should decrease and visa versa.

```{r}
cor(exercise_data$Exercise, exercise_data$BMI)
```

This shows that there is a negative correlation proving our initial hypothesis.

Now, lets visualize our data using a scatterplot.

```{r}
ggplot(data = exercise_data, mapping = aes(x = Exercise, y = BMI)) + 
  geom_point()
```

I was tricked! I see that the data points make a picture of a dinosaur.

This shows the importance of always looking at the data even if the original correlation makes sense.

## Question 2

I first installed "causact" in the console using the fucntion install.packages("causact").

```{r}
library(causact)
glimpse(corruptDF)
```

```{r}
?corruptDF
view(corruptDF)
```

CPI2017 is The Corruption Perceptions Index score for the year 2017. This index measures the perceived level of corruption of the public sector on a scale of 0-100. 0 means that a country is highly corrupt whereas 100 means that a country is not corrupt.

HDI2017 is the Human Development Index score for 2017. This index measures the achievement of human development across countries. It averages the simple average of a nation's longevity, education, and income.

## Question 3

Lets compare the relationship between HDI and CPI by creating a scatterplot.

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + 
  geom_point()

```

The graph shows that HDI and CPI have a positive relationship in 2017.

## Question 4

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + geom_smooth() +
  geom_point()
```

## Question 5

Lets explore this relationship by region. We can do this by using the "facet" function.

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region)
```

The trends of patterns are clear but some of the points/data seems cluttered. In order to make this more legible, we can reverse the scale of the x-axis.

## Question 6

I reversed the scale of the x-axis by using "scale_x_reverse" function.

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region) + scale_x_reverse()
```

## Question 7

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region) + scale_x_reverse() + labs(title = "The relationship between HDI and CPI", subtitle = "Each data point represents one country", caption = "Source: corruptDF")
```

I was able to add a title, subtitle, and caption to the plot by using the "labs" function.

## Question 8

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017, fill = region)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~region) + scale_x_reverse() + labs(title = "Relationship betweend HDI and CPI", subtitle = "Data points are countries", caption = "Source: corruptDF")

ggsave("HDI:CDI.pdf")
```

In order to save our data/plot, we can use the "ggsave" function
