---
title: "HW07"
author: "Kamran Acito"
format: html
editor: visual
embed-resources: true
---

```{r}
#| message: false
library(tidyverse)
library(here)
library(moderndive)
library(dplyr)
library(ggplot2)
theme_set(theme_minimal())

nobel_winners <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/nobel_winners.csv")
```

```{r}
nobel_winners_flt <- nobel_winners %>% 
  filter(category != "Peace") %>% 
  mutate(is_us = if_else(birth_country == "United States of America", 1, 0))
```

```{r}
true_prop <- nobel_winners_flt %>% 
  group_by(is_us) %>% 
  summarise(prop = n()/nrow(nobel_winners_flt))

ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = true_prop[2,2][[1]], linetype = "dashed")  + 
  labs(x = "Proportion")
```

```{r}
set.seed(33)
# Random sample 
our_sample <- sample_n(nobel_winners_flt, 
                       size = 25, 
                       replace = F)

sample_prop <- our_sample %>% 
  group_by(is_us) %>% 
  summarise(prop = n()/nrow(our_sample))

# How does our proportion look like? 
ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = true_prop[2,2][[1]], linetype = "dashed")  + 
  geom_vline(xintercept = sample_prop[2,2][[1]], linetype = "dotted", col = "red") +
  labs(x = "Proportion")
```

## Question 1

-   In this scenario, the scientific nobel prize winners are the population.

-   The sampling frame is a 25 person sample from the population.

-   The black line would represents the true proportion of Nobel Prize winners who were born in the US.

-   The red line represents the sampled proportion of winners.

## Question 2

We are going to take repeated samples of differing sizes taking 50 samples of 25 winners.

```{r}
Q2 <- nobel_winners_flt %>%
  rep_sample_n(size=25, reps=50)
prop_sample <- Q2 %>% 
  group_by(replicate) %>% 
  summarise(prop = sum(is_us)/25)
```

Now, lets plot the data including the mean proportion of the samples and the true proportion as calculated above.

```{r}
ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = mean(nobel_winners_flt$is_us), linetype = "dashed")  + 
  geom_vline(xintercept = mean(prop_sample$prop), linetype = "dotted", col = "red") +
  labs(x = "Proportion")
```

## Question 3

Now, we are going to take 50 samples of 100 samples each.

```{r}
Q3 <- nobel_winners_flt %>%
  rep_sample_n (size=100, reps=50)
prop_sample2 <- Q3 %>% 
  group_by(replicate) %>% 
  summarise(prop = sum(is_us)/100)
```

Lets plot the results

```{r}
ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = mean(nobel_winners_flt$is_us), linetype = "dashed")  + 
  geom_vline(xintercept = mean(prop_sample2$prop), linetype = "dotted", col = "red") +
  labs(x = "Proportion")
```

The bigger sample will have less variance because more data points leads to a more accurate estimate.

## Question 4

Now, we are going to calculate the standard deviation using sample sizes of 25 and 100.

```{r}
sd_25 <- sd(prop_sample$prop)

sd_25
```

```{r}
sd_100 <- sd(prop_sample2$prop)

sd_100
```

This does support my argument because the larger sample size has a smaller standard deviation meaning that there is less variance.

## Question 5

Now, we are going to repeat the processes above in order to find the proportion of women who have been awarded the Nobel Prize.

```{r}
np_women <- nobel_winners_flt %>%
   mutate(is_f = if_else(gender == "Female", 1, 0))
Q5 <- np_women %>%
  rep_sample_n(size=100, reps=50)
prop_sample3 <- Q5 %>% 
  group_by(replicate) %>% 
  summarise(prop = sum(is_f)/100)
prop_women <- np_women %>% 
  group_by(is_f) %>% 
  summarise(prop = n()/nrow(nobel_winners_flt))
```

```{r}
ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = prop_women[2,2][[1]], linetype = "dashed")  + 
  geom_vline(xintercept = prop_sample3[2,2][[1]], linetype = "dotted", col = "red") +
  labs(x = "Proportion")
```

## Question 6

From my understanding., the Central Limit Theorem states that a higher number of samples from a given population will lead to a more normal distribution. The samples will be closer to the mean creating less variance. Because we never know the true proportion of a population, we can make inferences/assumptions using the Central Limit Theorem.
